Normalization by Evaluation for Sized Dependent Types

### 2017-05-08 types17/types17.tex
[ TYPES 17 ] Proof reading.

\begin{abstract}
Sized types have been developed to make termination checking more perspicuous, more powerful, and more modular by integrating termination into type checking.  In dependently-typed proof assistants where proofs by induction are just recursive functional programs, the termination checker is an integral component of the trusted core, as validity of proofs depend on termination.  However, a rigorous integration of full-fledged sized types into dependent type theory is lacking so far.  Such an integration is non-trivial, as explicit sizes in proof terms might get in the way of equality checking, making terms appear distinct that should have the same semantics.
%
In this work, we integrate dependent types and sized types with
higher-rank size polymorphism, which is essential for generic
programming and abstraction.  We introduce a size quantifier $\forall$
which lets us ignore sizes in terms for equality checking, alongside
with a second quantifier $\Pi$ for abstracting over sizes that do
affect the semantics of types and terms.  Judgmental equality is
decided by an adaptation of normalization-by-evaluation. % for our new type theory, which features \emph{type shape}-directed reflection and reification.
\end{abstract}

@@

Agda \cite{agdawiki} features first class size polymorphism
\cite{abel:fics12} in contrast to theoretical accounts of sized
dependent types
\cite{bartheGregoirePastawski:lpar06,blanqui:rta04,sacchini:lics13}
who typically just have prenex (ML-style) size quantification.
Consequently, Agda's internal language contains size expressions in
terms wherever a size quantifier is instantiated.  However, these
size expressions, which are not unique due to subtyping, can get in
the way of reasoning about sizeful programs.
Consider the type of sized natural numbers.

>>>>

\begin{abstract}

Sized types have been developed to make termination checking more
perspicuous, more powerful, and more modular by integrating
termination into type checking.  In dependently-typed proof
assistants where proofs by induction are just recursive functional
programs, the termination checker is an integral component of the
trusted core, as validity of proofs depends on termination.
However, a rigorous integration of full-fledged sized types into
dependent type theory is lacking so far.  Such an integration is
non-trivial, as explicit sizes in proof terms might get in the way
of equality checking, making terms appear distinct that should have
the same semantics.
%
In this work, we integrate dependent types and sized types with
higher-rank size polymorphism, which is essential for generic
programming and abstraction.  We introduce a size quantifier $\forall$
which lets us ignore sizes in terms for equality checking, alongside
with a second quantifier $\Pi$ for abstracting over sizes that do
affect the semantics of types and terms.  Judgmental equality is
decided by an adaptation of normalization-by-evaluation.
% for our new type theory, which features \emph{type shape}-directed reflection and reification.
\end{abstract}

@@

Agda \cite{agdawiki} features first-class size polymorphism
\cite{abel:fics12} in contrast to theoretical accounts of sized
dependent types
\cite{bartheGregoirePastawski:lpar06,blanqui:rta04,sacchini:lics13}
who typically just have prenex (ML-style) size quantification.
Consequently, Agda's internal language contains size expressions in
terms wherever a size quantifier is instantiated.  However, these
size expressions, which are not unique due to subtyping, can get in
the way of reasoning about sizeful programs.
Consider the type of sized natural numbers.
